Manuscript draft

### 2019-08-22 manuscript/manuscript draft.Rmd
updated manuscript draft

Globally, forests play an important role in the carbon cycle and are an important component of global carbon dioxide budgets [@luyssaert_old-growth_2008]. They show higher levels of productivity than non-forest  terrestrial ecosystems [@del_grosso_global_2008], and as a result achieve significant carbon sequestration and storage. Estimating the total role of forests in the carbon cycle is challenging, but studies indicate that old growth forests alone sequester up to 1.4 GtCyr^-1^ [@malhi_carbon_1999], while the total sequestration of carbon by established forests globally could be up to 2.4 GtCyr^-1^, with the largest sinks being in old-growth tropical forests [@pan_large_2011]. As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change. Understanding the patterns of forest productivity on a global scale, and the drivers behind them, is therefore a priority in forest research.

@@

Table: Definitions of variables used in analysis

@@

*Climate datasets.* Where site-level data on mean annual temperature, mean annual precipitation, and latitude were available in the primary literature, this data was compiled and entered directly into the ForC database. In addition to this data, climate data for each site was extracted from five open-access climate datasets based on site geographic co-ordinates.

@@

@@

>>>>

Globally, forests play an important role in the carbon cycle and are a major component of global carbon dioxide budgets [@luyssaert_old-growth_2008]. They show higher levels of productivity than non-forest terrestrial ecosystems [@del_grosso_global_2008], and as a result achieve significant carbon sequestration and storage. Estimating the total role of forests in the carbon cycle is challenging, but studies indicate that old growth forests alone sequester up to 1.4 GtCyr^-1^ [@malhi_carbon_1999], while the total sequestration of carbon by established forests globally could be up to 2.4 GtCyr^-1^, with the largest sinks being in old-growth tropical forests [@pan_large_2011]. As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change. Understanding the patterns of forest productivity on a global scale, and the drivers behind them, is therefore a priority in forest research.

@@

Table: Definitions and sample sizes of variables used in analysis

@@

*Climate datasets.* Where site-level data on mean annual temperature, mean annual precipitation, and latitude were available in the primary literature, this data was compiled and entered directly into the ForC database. In addition to this data, climate data for each site was extracted from five open-access climate datasets based on site geographic co-ordinates. Where site-level data was missing for mean annual temperature and/or mean annual precipitation, data was extracted from the WorldClim dataset.

@@

Additionally, two climate variables were derived from the above datasets: maximum vapour pressure deficit, defined as the vapour pressure deficit of the month with the largest deficit; and water stress months, defined as the number of months annually where precipitation was lower than potential evapotranspiration.

*Length of growing season.* Growing season months were defined as months with mean minimum temperature > 0.5. Growing season months were initially calculated following methods used by Kerkhoff et al. (2005), which additionally required that growing season months had a moisture index, defined as (MAT - PET)/PET, > -0.95. Michaletz et al. (2014) included an equivalent requirement in their calculation of growing season length. However, we found that including this requirement had no effect on the estimates of growing season length, and so chose to exclude it.

Monthly data for PET, precipitation, and temperature was downloaded from the Climate Research Unit (CRU) time-series dataset v 4.03 [@harris_updated_2014], and for solar radiation from WorldClim2 [@fick_worldclim_2017], and used to calculate mean monthly PET, precipitation, temperature and solar radiation during the growing season. Total growing season precipitation and solar radiation were also calculated.

@@

Models were run for total annual productivity against annual climate variables, and for monthly growing season productivity, defined as total productivity/length of growing season, against growing season climate variables.
